Mechanized Reasoning about a Capability Machine

### 2019-10-17 prisc-2020/paper.tex
types and commenting out a technical paragraph in section 3

\begin{itemize}
\item we define the operational semantics of the CM
\item we prove basic hoare triples for each of the atomic instructions
  of the CM (i.e., for load, store, etc.)
\item non-trivial
  \begin{itemize}
  \item while Iris is a framework and supports many languages, it is
    geared towards models of higher-level languages which abstract from
    program counters and the fact that programs are stored in memory.
    -- here our low-level machine model includes registers, in particular
    the program counter register, and instructions are stored in memory.
  \item to capture the semantics of the CM in the Iris framework, 
    we introduce an abstract notion of instruction, whose
    operational meaning is to execute the next instruction the program counter
    points to, and abstract notions of values, failed and halted.
  \item for each machine instruction, we can then prove a hoare triple
    which expresses the pre- and post-conditions for executing the instruction
  \end{itemize}
\end{itemize}

@@

To capture the semantics of the Capability Machine in the Iris framework, we introduce an abstract notion of an instruction, whose operational meaning is to execute the instruction the program counter points to, and an abstraction notion of values for halted and failed configurations.

@@

\subsection{The Value Relation}
\begin{itemize}
	\item definition of the relation and World
	\item LS definitions
	\item now: standard islands are not used 
\end{itemize}

@@

\subsection{Reasoning about Local State}
\begin{itemize}
	\item How to reason about local state in high level languages
	\item public/private distinction
	\item TT style
	\item difficulties w/ iris invariants 
\end{itemize}

@@

In our case however, calls are not always well-bracketed. \citeauthor{skorstengaardESOP18}. distinguishes between well-bracketed and non well-bracketed calls by respectively using public- and private future world relations given the appropriate conditions.

Iris invariants do not make such a distinction. Using invariants alone is therefore not sufficient for our purposes. Instead we explicitly define notions of public and private future world relations, and use these notions to distinguish between well-bracketed and non well-bracketed calls.

A callback is an executable capability. This distinction must therefore be done by the value relation. However, if the value relation is indexed by a world, we again have difficulties using Iris invariants. Asserting $\exists w, \mathcal{V}(W)(w)$ as invariant does not allow for $W$ to change.

The difficulty is to support reasoning about local state, while distinguishing between public- and private future world relations in the value relation, rather than the expression relation. We implement this by using Iris' saved predicates, which allows us to save a predicate by associating it to a unique name. Using that unique name, we are then able to refer to that predicate somewhere else, and state it with the argument that we want. That way, if we save the value relation predicate and associate it to a location, we can later on state that predicate and apply it to the future world in question.

>>>>

%\begin{itemize}
%\item we define the operational semantics of the CM
%\item we prove basic hoare triples for each of the atomic instructions
%  of the CM (i.e., for load, store, etc.)
%\item non-trivial
%  \begin{itemize}
%  \item while Iris is a framework and supports many languages, it is
%    geared towards models of higher-level languages which abstract from
%    program counters and the fact that programs are stored in memory.
%    -- here our low-level machine model includes registers, in particular
%    the program counter register, and instructions are stored in memory.
%  \item to capture the semantics of the CM in the Iris framework, 
%    we introduce an abstract notion of instruction, whose
%    operational meaning is to execute the next instruction the program counter
%    points to, and abstract notions of values, failed and halted.
%  \item for each machine instruction, we can then prove a hoare triple
%    which expresses the pre- and post-conditions for executing the instruction
%  \end{itemize}
%\end{itemize}

@@

To capture the semantics of the Capability Machine in the Iris framework, we introduce an abstract notion of an instruction, whose operational meaning is to execute the instruction the program counter points to, and an abstract notion of values for halted and failed configurations.

@@

\subsection{The Value Relation}
%\begin{itemize}
%	\item definition of the relation and World
%	\item LS definitions
%	\item now: standard islands are not used 
%\end{itemize}

@@

\subsection{Reasoning about Local State}
%\begin{itemize}
%	\item How to reason about local state in high level languages
%	\item public/private distinction
%	\item TT style
%	\item difficulties w/ iris invariants 
%\end{itemize}

@@

In our case however, calls are not always well-bracketed. \citeauthor{skorstengaardESOP18} distinguishes between well-bracketed and non well-bracketed calls by respectively using public- and private future world relations at the site of the callback.

%Iris invariants do not make such a distinction. Using invariants alone is therefore not sufficient for our purposes. Instead we explicitly define notions of public and private future world relations, and use these notions to distinguish between well-bracketed and non well-bracketed calls.

A callback is an executable capability. This distinction must therefore be done by the value relation. However, if the value relation is parametrized by a world, we have difficulties using Iris invariants. Asserting $\mathcal{V}(W)(w)$ as invariant does not allow for $W$ to change.

The difficulty lies in supporting reasoning about local state, while distinguishing between public- and private future world relations in the value relation, rather than the expression relation. We implement this by using Iris' saved predicates, which allows us to save a predicate by associating it to a unique name. Using that unique name, we are then able to refer to that predicate somewhere else, and state it with the argument that we want. That way, if we save the value relation predicate and associate it to a location, we can later on state that predicate and apply it to the future world in question.
